Spatial networks
========================================================
A missing sp-compatible spatial data class is that for spatial networks. This document will do some simple things by combining the spatial (`SpatialLines`) classes of `sp` with the graph classes and methods from `igraph`. We will compute and plot the shortest path through a graph made of Delauney triangle edges from random points.

To start, we'll need to inport the spatial and graph packages:

```{r}
library(sp)
library(igraph)
```
and define the igraph class as S4, in order to reuse it in a new S4 class
```{r}
setClass("igraph")
```
Next, we define the `SpatialLinesNetwork` class, consisting of 
- a `SpatialLines object`, containing the spatial lines representing the edges
- an `igraph` object, containing the network topology
- a neighbourhood list `nb`, containing for each vertex which edges it connets to:
```{r}
setClass("SpatialLinesNetwork",
    representation(sl = "SpatialLines", g = "igraph", nb = "list"),
    validity = function(object) {
        stopifnot(length(object@sl) == length(E(object@g)))
        stopifnot(length(object@nb) == length(V(object@g)))
    }
)
```
The following section creates a `SpatialLines` object, and converts it into a `SpatialLinesNetwork`. As you can see, it only accepts "simple" objects, having only one `Line` per `Lines` set. Much of what it does is figure out the common points (igraph: vertices) where line segments (igraph: edges) meet.
```{r}
SpatialLinesNetwork = function(sl) {
    stopifnot(is(sl, "SpatialLines"))
    if (!is(sl, "SpatialLinesDataFrame"))
        sl = new("SpatialLinesDataFrame", sl,
            data = data.frame(id=1:length(sl)))
    if(!all(sapply(sl@lines, length) == 1))
        stop("SpatialLines is not simple: each Lines element should have only a single Line")
    startEndPoints = function(x) {
        firstLast = function(L) {
            cc = coordinates(L)[[1]]
            rbind(cc[1,], cc[nrow(cc),])
        }
        do.call(rbind, lapply(x@lines, firstLast))
    }
    s = startEndPoints(sl)
    zd = zerodist(SpatialPoints(s))
    pts = 1:nrow(s)

    # the following can't be done vector-wise, there is a progressive effect:
    if (nrow(zd) > 0) {
        for (i in 1:nrow(zd))
            pts[zd[i,2]] = pts[zd[i,1]]
    }
    stopifnot(identical(s, s[pts,]))

    # map to 1:length(unique(pts))
    pts0 = match(pts, unique(pts))
    node = rep(1:length(sl), each = 2)
    nb = lapply(1:length(unique(pts)), function(x) node[which(pts0 == x)])
    g = graph(pts0, directed = FALSE) # edges
    nodes = s[unique(pts),]
    g$x = nodes[,1] # x-coordinate vertex
    g$y = nodes[,2] # y-coordinate vertex
    g$n = as.vector(table(pts0)) # nr of edges
    # line lengths:
    sl$length = sapply(sl@lines, function(x) LineLength(x@Lines[[1]]))
    E(g)$weight = sl$length
    # create list with vertices, starting/stopping for each edge?
    # add for each SpatialLines, the start and stop vertex
    pts2 = matrix(pts0, ncol = 2, byrow = TRUE)
    sl$start = pts2[,1]
    sl$end = pts2[,2]
    new("SpatialLinesNetwork", sl = sl, g = g, nb = nb)
}
```

```{r}
l0 = cbind(c(1,2), c(0,0))
l1 = cbind(c(0,0,0),c(0,1,2))
l2 = cbind(c(0,0,0),c(2,3,4))
l3 = cbind(c(0,1,2),c(2,2,3))
l4 = cbind(c(0,1,2),c(4,4,3))
l5 = cbind(c(2,2), c(0,3))
l6 = cbind(c(2,3), c(3,4))
l = list(
    Lines(list(Line(l0)), "e"),
    Lines(list(Line(l1)), "a"),
    Lines(list(Line(l2)), "b"),
    Lines(list(Line(l3)), "c"),
    Lines(list(Line(l4)), "d"),
    Lines(list(Line(l5)), "f"),
    Lines(list(Line(l6)), "g")
)
sl = SpatialLines(l)
sln = SpatialLinesNetwork(sl)
```
We can plot these data in geographical space, using colour to denote the number of edges at each vertex:
```{r  fig.width=7, fig.height=6}
plot(sln@g$x, sln@g$y, col = sln@g$n, pch=16,cex=2, asp = 1)
lines(sl)
text(sln@g$x, sln@g$y, E(sln@g),pos=4)
```
The plot in the graph space looks like this:
```{r}
plot(sln@g)
```

A larger example: shortest path through a Delauny triangulations network
------------------------------------------
We'll use Delauny triangulation from package `deldir`:
```{r}
require(deldir)
```
The following function converts a set of points into a `SpatialPolygons` or into a `SpatialLines` object:
```{r}
dd <- function(x, ..., to = "polygons") { 
    stopifnot(is(x, "Spatial"))
    cc = coordinates(x)
    dd = deldir(list(x = cc[,1], y = cc[,2]),...)
    if (to == "polygons") {
        tm = triMat(dd) 
        fn = function(ix) {
            pts = tm[ix,]
            pts = c(pts, pts[1])
            Polygons(list(Polygon(rbind(cc[pts,]))), ix)
        }
        SpatialPolygons(lapply(1:nrow(tm), fn), proj4string = x@proj4string)
    } else if (to == "lines") {
        segs = dd$delsgs
        lst = vector("list", nrow(segs))
        for (i in 1:nrow(segs))
            lst[i] = Lines(list(Line(cc[c(segs[i,5], segs[i,6]),])), i)
        SpatialLines(lst, proj4string = x@proj4string)
    } else stop("argument to should be polygons or lines")
}   
```
We'll generate a set of 100 points in a unit square:
```{r}
set.seed(5432)
x = runif(100)
x = x[order(x)]
y = runif(100)
pts = SpatialPoints(cbind(x,y))
```
Next, we'll get the `SpatialLines` object from it:
```{r}
sl = dd(pts, to = "lines")
```
and plot it:
```{r}
plot(sl)
```
From this object, we can create a `SpatialLinesNetwork` by
```{r}
ln = SpatialLinesNetwork(sl)
```
when plotted, we can again use colour to denote the number of vertices connected to each edge:
```{r fig.width=7, fig.height=6}
plot(ln@g$x, ln@g$y, col = ln@g$n, pch = 16, cex = 2, asp = 1)
lines(sl, col = 'grey')
```
Now we compute the shortest path from the left-most point (1) to the right-most one (100):
```{r}
sp = get.shortest.paths(ln@g, 1, 100)[[1]]
```
and plot it
```{r fig.width=7, fig.height=7}
plot(ln@g$x, ln@g$y, col = ln@g$n, pch = 16, cex = 1.5, asp = 1)
lines(sl, col = 'grey')
points(ln@g$x[sp], ln@g$y[sp], col = 'red', cex = 2)
text(ln@g$x[c(1,100)], ln@g$y[c(1,100)], c("start", "end"), pos = 4)
```
As the edge weights are computed by `Line` lengths, this is the geographically shortest path.

I haven't figure out how to draw the shortest path by selecting the appropriate line segments; I guess I have to use
```{r}
sp
```
which gives the node sequence, and match each pair, `r sp[1]` - `r sp[2]`, `r sp[2]` - `r sp[3]` etc., to the start and end fields (which are now un-sorted!) -- some challenges ahead.