Trajectories represented by a lists of lists
============================
Along the lines of classes `SpatialPolygonsDataFrame` and `SpatialLinesDataFrame`
in package `sp`, we will implement Trajectories here as (typed) lists of lists.
The reason for this is that trajectory data sets usually contain sets of tracks,
or trips, per person (or animal, or object), and then groups these.

Sources of this document are found at <http://github.com/edzer/rpubs>.

The class will be demonstrated with a small excerpt of the geolife data.
```{r}
library(sp)
library(spacetime)
```

The first class is `Track`, and contains a single track, or trip, followed by a person, animal or object. This means that consecutive location/time stamps are not interrupted by a period of substantially other activity.
```{r}
Track = setClass("Track",
  contains = "STIDF", # locations, times and attribute data about the points
  representation(connections = "data.frame"), # attribute data BETWEEN points: speed, direction etc.
  validity = function(object) {
	stopifnot(nrow(object@connections) + 1 == nrow(object@data))
    return(TRUE)
  }
)
```
This object extends `STIDF`, where locations and times as well as attributes measured at these locations and times, such as elevation, are stored. The function `Track` can now be used to build such an object from its components. The slot `connections contains data about the segments between consecutive ST points.

A further utility function, `CreateTrack` does a bit more work and computes segment lengths and speed, from an `STIDF` object, and puts it in the `connections` slot:
```{r}
CreateTrack = function(obj, df = NULL) { # computes segment lenghts
	cc = coordinates(obj@sp)
	ll = is.projected(obj@sp)
	if (is.na(ll))
		ll = FALSE
	L = LineLength(cc, sum = FALSE, longlat = ll)
	if (is.null(df))
		df = data.frame(lengths = lengths)
	else {
		stopifnot(nrow(df) == length(L))
		df$lengths = L
	}
	#df$speed = df$lengths / as.numeric(diff(index(obj@time)))
  df$speed = df$lengths / diff(as.numeric(index(obj@time)))
	Track(obj, connections = df)
}
```

The next class, `Tracks`, takes a collection (list) of `Track` objects, along with a metadata data.frame `tracksData` containing (summary) information, one record per `Track` (e.g. total length, duration, average speed). It is assumed that each `Tracks` object refers to a single person, animal or object.
```{r}
Tracks = setClass("Tracks", # a collection of Track objects for single ID (car, person etc)
	representation(tracks = "list", tracksData = "data.frame"),
	validity = function(object) {
		stopifnot(all(sapply(object@tracks, function(x) is(x, "Track"))))
		stopifnot(nrow(object@tracksData) == length(object@tracks))
		stopifnot(length(object@tracks) > 0)
		return(TRUE)
	}
)
```

`TracksCollection` finally collects `Tracks` collections for several persons, animals or objects, and has a slot `tracksCollectionData` with a summary record for each person/animal/object:
```{r}
TracksCollection = setClass("TracksCollection", # collection of Tracks for several IDs 
	representation(tracksCollection = "list", tracksCollectionData = "data.frame"),
	validity = function(object) {
		stopifnot(all(sapply(object@tracksCollection, class) == "Tracks"))
		stopifnot(length(object@tracksCollection) == 
			nrow(object@tracksCollectionData))
		stopifnot(length(object@tracksCollection) > 0)
		names = names(object@tracksCollection)
		stopifnot(!(is.null(names) || any(is.na(names))))
		return(TRUE)
	}
)
```

Next, we will read a number of geolife trajectories, for the following persons:
```{r}
IDS = c("079", "095", "111", "127", "143", "159", "175")
```


```{r}
# Geolife trajectories,
# documentation: http://research.microsoft.com/apps/pubs/?id=152176 :
# data: http://research.microsoft.com/en-us/projects/geolife/
# or http://ftp.research.microsoft.com/downloads/b16d359d-d164-469e-9fd4-daa38f2b2e13/Geolife%20Trajectories%201.2.zip
setwd("/home/edzer/Downloads/Geolife Trajectories 1.3/Data/")
#sel = 1:2
#sel = TRUE
i = j = 1
#dirs = list.files("Data")[sel]
crs = CRS("+proj=longlat +datum=WGS84")
#pb = txtProgressBar(style = 3, max = length(IDS))
elev = numeric(0)
lst0 = list()
for (d in IDS) {
	dir = paste(d, "Trajectory", sep = "/")
	print(dir)
	lst = list()
	files = list.files(dir, pattern = "*plt", full.names = TRUE)
	i = 1
	for (f in files) {
		tab = read.csv(f, skip = 6, stringsAsFactors=FALSE)
		tab$time = as.POSIXct(paste(tab[,6],tab[,7]))
		tab[tab[,4] == -777, 4] = NA # altitude 
		tab = tab[,-c(3,5,6,7)]
		names(tab) = c("lat", "long", "elev", "time")
		if (all(tab$lat > -90 & tab$lat < 90 & tab$long < 360 
				& tab$long > -180)) {
			stidf = STIDF(SpatialPoints(tab[,2:1], crs), tab$time, tab)
			conn = tab[-1,"elev",drop=FALSE]
			lst[[i]] = CreateTrack(stidf, conn)
			i = i+1
		}
	}
	names(lst) = f
	n = sapply(lst, length)
	lst0[[j]] = Tracks(tracks = lst, tracksData = data.frame(length = n))
	#setTxtProgressBar(pb, j)
	j = j+1
}
names(lst0) = IDS
df = data.frame(IDS=IDS)
TR = TracksCollection(tracksCollection = lst0, tracksCollectionData = df)
object.size(TR)
```

Next, we define a few utility functions to convert the track data into `data.frame` objects, with `NA` separated trips:
```{r}
setAs("Track", "data.frame", function(from)
	as(as(from, "STIDF"), "data.frame")
)
setAs("Tracks", "data.frame", function(from)
	do.call(rbind, lapply(from@tracks, function(x) rbind(as(x, "data.frame"), NA)))
)
setAs("TracksCollection", "data.frame", function(from) {
		l = lapply(from@tracksCollection, function(x) rbind(as(x, "data.frame"), NA))
		n = sapply(l, nrow)
		ret = do.call(rbind, l)
		data.frame(ret, IDs = rep(names(from@tracksCollection), times = n))
	}
)
#as(TR@tracksCollection[[1]]@tracks[[1]], "data.frame")
#as(TR@tracksCollection[[1]], "data.frame")
```
and we use this:
```{r}
df = as(TR, "data.frame")
# focus on Bejing:
```
then, we can plot the trajectories read:
```{r}
setMethod("plot", "TracksCollection", function(x, ..., type = 'l', colorBy = "IDs") {
		df = as(x, "data.frame")
		if (colorBy == "IDs")
			col = as.numeric(as.factor(df$IDs))
		# print(col)
		plot(lat ~ long, df, asp = 1, type = type, col = col, ...)
	}
)
#plot(TR, xlim=c(116.3,116.5),ylim=c(39.8,40))
plot(TR)
```

```{r}
library(lattice)
xyplot(lat~long, df, type='l',
       xlim=c(116,116.5),
       ylim=c(39.5,40.5),
       groups=IDs,
       col=1:20)
```

```{r}
setMethod("stplot", "TracksCollection", function(obj, ..., mode = "byID") {
    sp = obj@tracksCollection[[1]]@tracks[[1]]@sp
    if (mode == "byID")
		  xyplot(lat~long, as(obj, "data.frame"), type='l', 
		  	groups = IDs, asp = mapasp(sp), ...)
    else
        stop("no other modes implemented")
	}
)
stplot(TR, xlim=c(116.3,116.5),ylim=c(39.8,40), col = 1:20)
```
